---
title: "My Website"
description: |
  Welcome to the website. I hope you enjoy it!
output: bookdown::gitbook
site: bookdown::bookdown_site
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

#site: distill::distill_website
#output: distill::distill_article

# output: bookdown::gitbook
# site: bookdown::bookdown_site

```


<!--chapter:end:index.Rmd-->

---
title: "About this site"
description: |
  Some additional details about the website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


<!--chapter:end:about.Rmd-->

---
title: "Saegus Interview Project: Predicting Bike Rentals"
author:
  - name: "Joffrey JOUMAA"
    url: https://github.com/SESjo
    affiliation: TheGreenData
    affiliation_url: https://www.thegreendata.com
description: | 
  Scientific and technical writing, native to the web
output: 
  bookdown::html_document2:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

<style>
body {
text-align: justify}
</style>

```{r setup, include = FALSE}
# options
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      comment = NA, 
                      cache = FALSE,
                      fig.align = "center"
)

knitr::opts_knit$set(root.dir = "~/Documents/post/saegus/bike_rental_hour/")


    # code_folding: hide
    # df_print: paged
    # fig_caption: yes
    # toc: yes
    # toc_float:
    #   collapsed: yes
    #   smooth_scroll: no
```

# Context

## Overview

Many American cities have communal bike sharing stations where you can rent bicycles by the hour or day. Washington, D.C. is one of these cities. The District collects detailed data on the number of bicycles people rent by the hour and day.

[Hadi Fanaee-T](http://www.liaad.up.pt/area/fanaee) at [the University of Porto](http://www.up.pt/) compiled this data into a CSV file, which you'll be working with in this project. The file contains `17380` rows, with each row representing the number of bike rentals for a single hour of a single day. You can download the data from [the University of California](http://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset), Irvine's website.

Here's what the first five rows look like:
```{r loadDataset}
# packages
library(data.table)
library(GGally)
library(caret)
library(ggplot2)
library(doMC)
library(randomForest)
library(corrplot)
library(plotly)

# loading file
dataset = fread("bike_rental_hour.csv")

# print first rows
head(dataset)
```

Here are the descriptions for the relevant columns:

- `instant` - A unique sequential ID number for each row
- `dteday` - The date of the rentals
- `season` - The season in which the rentals occurred
- `yr` - The year the rentals occurred
- `mnth` - The month the rentals occurred
- `hr` - The hour the rentals occurred
- `holiday` - Whether or not the day was a holiday
- `weekday` - The day of the week (as a number, `0` to `7`)
- `workingday` - Whether or not the day was a working day
- `weathersit` - The weather (as a categorical variable)
- `temp` - The temperature, on a `0-1` scale
- `atemp` - The adjusted temperature
- `hum` - The humidity, on a `0-1` scale
- `windspeed` - The wind speed, on a `0-1` scale
- `casual` - The number of casual riders (people who hadn't previously signed up with the bike sharing program)
- `registered` - The number of registered riders (people who had already signed up)
- `cnt` - The total number of bike rentals (`casual` + `registered`)

## Instructions

Let's say your customer want to predict the total number of bikes people rented in a given hour (`cnt` column ).

In this project you'll have to provide a clear and meaningfull data analysis using data story tellings approach and statistical technics.


Have Fun!

# Data Pretreatment

For this part, I've simply checked the presence of missing value.

```{r naCheckCode}
# the coding way 
dataset[, any(is.na(.SD))]
```

```{r naCheckGraph, f}
# the graphical way
dataPlot = melt(is.na(dataset))

ggplot(dataPlot, aes(x = Var2, y = Var1, fill = value))+
  geom_tile()+
  labs(x = "Attributes", y = "Rows")+
  scale_fill_manual(values = c("white", "black"), 
                    labels = c("Real", "Missing")) +
  theme(legend.position = "top", 
        axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.key = element_rect(colour = "black"))
```

Which is not the case.

I've also converted few numerical attributes as factors to ease data visualization.

```{r dataConvert}
# date format
dataset[, dteday := as.POSIXct(format(dteday,                 
                                      format = "%Y-%m-%d",
                                      tz = "CET"))]

# factors
dataset[, ':=' (season = as.factor(season),                         
                holiday = as.factor(holiday),
                mnth = as.factor(mnth),
                hr = as.factor(hr),
                yr = as.factor(yr),
                weekday = as.factor(weekday),
                workingday = as.factor(workingday),
                weathersit = as.factor(weathersit))]
```

<!--chapter:end:presentation_jjoumaa.Rmd-->

